### 2023-10-28 recs-abund-yield-parameters.qmd
initial commit

---
title: "Recruits, abundance, and yield time series and contours"
author: "Botsford, Hastings, White, and Kilduff"
execute: 
  warning: false
format:
  html: 
    page-layout: full
    embed-resources: true
    anchor-sections: true
    fig-width: 7
    fig-height: 5
  pdf:
    fig-width: 7
    fig-height: 5
  docx:
    fig-width: 7
    fig-height: 5
editor: visual
params:
  experiment: "larval_pool_blue_rockfish_mei_135_75"
---

```{r}
#| label: Extract experiemental parameters
#| echo: false

exp <- params$experiment

valid_dispersal <- c("larval_pool", "no_dispersal")
valid_species <- c("blue_rockfish", "cabezon", "china_rockfish", "pacific_cod", "kelp_bass")
valid_noise <- c("mei_135_75", "white_135_75")

current_dispersal <- valid_dispersal[stringr::str_detect(exp, valid_dispersal)]
current_species <- valid_species[stringr::str_detect(exp, valid_species)]
current_noise <- valid_noise[stringr::str_detect(exp, valid_noise)]

pretty_dispersal <- c("Larval pool", "No dispersal")
pretty_species <- c("Blue rockfish", "Cabezon", "China rockfish", "Pacific cod", "Kelp bass")
pretty_noise <- c("ENSO", "White")

nice_dispersal <- pretty_dispersal[stringr::str_detect(exp, valid_dispersal)]
nice_species <- pretty_species[stringr::str_detect(exp, valid_species)]
nice_noise <- pretty_noise[stringr::str_detect(exp, valid_noise)]

out_path <- file.path("output", current_species)

if (!dir.exists(out_path)) {
  dir.create(out_path, recursive = TRUE) 
}

```

### Experiment: `r params$experimenft`

This is a parameterized report for:

-   `r nice_species`
-   `r nice_noise`
-   `r nice_dispersal`

```{r}
#| label: load-packages-data
#| echo: false

library(targets)
suppressMessages(library(tidyverse))
library(duckdb)
options(dplyr.summarise.inform = FALSE)
# Load relevant data from target objexts
tar_load(sim_len)
tar_load(species_flep_fished_ind)
tar_load(fishing_mortality_values)
tar_load(species_parms)
tar_load(sim_species_parms)
tar_load(sim_species_derived_vars)

# Load project functions
source("R/functions.R")

# constants
burn_in_parm <- 500
sim_len_parm <- 135

x_text_position_parm <- 550

waa_g <- sim_species_parms[[nice_species]][["biom_const"]] * sim_species_derived_vars[[nice_species]][["length_at_age"]] ^
                    sim_species_parms[[nice_species]][["biom_exp"]]

# use functions to create useful data structures
# FLEP to F reference tables by species
spec_fleps_2_fs <-
  get_fs_from_fleps(species = nice_species,
                    flep_inds = species_flep_fished_ind,
                    f_vals =  fishing_mortality_values)

```

```{r}
#| label: functions
#| echo: false

# query specified simulation results and aggregate yield or recruits for subsequent analyses
qry_results_table_agg <- function(ddb_table,
                                  metric = c("recruits", "yield", "abundance"),
                                  burn_in = 500,
                                  sim_len = 135,
                                  reserve_fracs = NULL,
                                  flep_ratios = NULL,
                                  dir_duckdb = "data/sim_out.duckdb") {
  sim_yrs <- burn_in + sim_len
  age_filter <- ifelse(metric == "recruits", "age = 1 and ", "")

  reserve_fracs_qry <- ifelse(length(reserve_fracs) > 0,
                              paste0(" and cast(reserve_frac as decimal(5, 2)) in (", stringr::str_flatten_comma(format(reserve_fracs,  nsmall = 2)), ")"),
                              "")
  flep_ratios_qry <- ifelse(length(flep_ratios) > 0, 
                           paste0(" and cast(reserve_frac as decimal(5, 2)) in (", stringr::str_flatten_comma(format(flep_ratios, nsmall = 2)), ")"),
                           "")

  agg_qry <-
    paste0(
      "SELECT experiment, noise, species, bh_alpha, rec_sd, reserve_frac,
             flep_ratio, sim_num, year, age, sum(",
      ifelse(metric == "yield", "yield", "number"), ") as ",  metric,
      " FROM '", ddb_table,
      "' WHERE ", age_filter, "year > ", burn_in,
      " and year <= ", sim_yrs,
      reserve_fracs_qry, 
      flep_ratios_qry,
      " GROUP BY experiment, noise, species, bh_alpha, rec_sd, reserve_frac, flep_ratio, sim_num, year, age
        ORDER BY experiment, noise, species, bh_alpha, rec_sd, reserve_frac, flep_ratio, sim_num, year, age"
    )
  con <- DBI::dbConnect(duckdb::duckdb(),
                      dbdir = dir_duckdb,
                      read_only = FALSE)
  agg_df <- dbGetQuery(con, agg_qry)
  DBI::dbDisconnect(con, shutdown = TRUE)
  return(dplyr::tibble(agg_df))
}

make_df_yield_exp_spec_noise_rec_fs <-
  function(df_exp_spec_noise_rec,
           fleps_2_fs,
           f_maxes_yield) {
    df_exp_spec_noise_rec_fs <- df_exp_spec_noise_rec |>
      left_join(.,
                y = fleps_2_fs,
                by = join_by(flep_ratio_ch == flep_vals_ch)) %>%
      dplyr::mutate(f_fmsy = round(f_vals / f_maxes_yield, 2))
    return(df_exp_spec_noise_rec_fs)
  }

get_exp_details <- function(df_after_qry_results_table_agg) {
  exp_details <- df_after_qry_results_table_agg %>%
  summarise(experiment = unique(experiment), 
            noise = unique(noise),
            species = unique(species),
            bh_alpha = round(unique(bh_alpha), 4),
            rec_sd = unique(rec_sd))
  return(exp_details)
}

get_f_maxes_yield <- function(yield_disp_sp_noise, 
                              fleps_2_fs) {
  # Find the fishing mortality level that maximizes yield
  # with no reserves
  F_maxes_yield <- yield_disp_sp_noise %>%
  dplyr::filter(near(reserve_frac, 0)) %>%
  dplyr::left_join(., y = fleps_2_fs, by = join_by(flep_ratio_ch == flep_vals_ch)) %>%
  dplyr::group_by(f_vals, flep_ratio) %>%
  dplyr::summarise(yields = sum(yield)) %>%
  dplyr::ungroup() %>%
  dplyr::filter(yields == max(yields)) %>%
  .$f_vals

  return(F_maxes_yield[[1]])
}

# Time series prep tplotting function
make_df_exp_spec_noise_rec_fs <- function(df_exp_spec_noise_rec,
                                          fleps_2_fs,
                                          f_maxes_yield) {
  df_exp_spec_noise_rec_fs <- df_exp_spec_noise_rec %>%
    mutate(flep_ratio_ch = as.character(flep_ratio)) %>%
      left_join(., y = fleps_2_fs, by = join_by(flep_ratio_ch == flep_vals_ch)) %>%
      dplyr::mutate(f_fmsy = round(f_vals / f_maxes_yield, 2)) %>%
    select(-flep_vals)
  return(df_exp_spec_noise_rec_fs)
}

prep_ann_yield_for_plot <- function(df_exp_spec_noise_rec_fs,
                                    frac_reserves = c(0, 0.25, 0.50),
                                    f_fmsys = c(0.52, 1.00, 2.09)) {

  if (length(f_fmsys) < 1 | is.null(f_fmsys)) {
    stop("Please specify f_fmsy values to use")
  }

  if (length(frac_reserves) < 1 | is.null(frac_reserves)) {
    stop("Please specify frac_reserves values to use")
  }

  # make a sub function
    med_yield <- df_exp_spec_noise_rec_fs %>%
      dplyr::filter(near(f_fmsy, 1), near(reserve_frac, 0))  %>%
      dplyr::group_by(
        experiment,
        noise,
        species,
        bh_alpha,
        rec_sd,
        reserve_frac,
        flep_ratio,
        sim_num,
        year
      ) %>%
      dplyr::summarise(ann_yield = sum(yield)) %>%
      dplyr::ungroup() %>%
      dplyr::summarise(med_yield_f_fmsy = median(ann_yield)) %>%
      .$med_yield_f_fmsy

    # make a sub function
    main_plt_df <- df_exp_spec_noise_rec_fs %>%
      dplyr::filter(reduce(map(
        frac_reserves, near, x = reserve_frac, tol = 1e-4
      ), `|`)) %>%
      dplyr::group_by(reserve_frac, flep_ratio, f_fmsy, year) %>%
      dplyr::summarize(annual_yield = sum(yield)) %>%
      dplyr::ungroup() %>%
      dplyr::filter(reduce(map(
        f_fmsys, near, x = f_fmsy, tol = 1e-4
      ), `|`))

    # make a sub function
    supp_plt_df <- main_plt_df %>%
      dplyr::mutate(below_thresh = if_else(annual_yield <= med_yield, 1, 0)) %>%
      dplyr::group_by(reserve_frac, f_fmsy) %>%
      dplyr::summarize(
        mean_yield = round(mean(annual_yield, na.rm = TRUE), 2),
        mean_yield_lab = paste0("Mean yield: ", mean_yield),
        sd_yield = round(sd(annual_yield, na.rm = TRUE), 2),
        sd_yield_lab = paste0("SD yield: ", sd_yield),
        cv_yield = round(sd(annual_yield, na.rm = TRUE)/mean(annual_yield, na.rm = TRUE), 2),
        cv_yield_lab = paste0("CV yield: ", cv_yield),
        yrs_below_thresh = paste("% YR below thresh: ", round(sum(below_thresh)/135, 2))
      )
    return(list(main_plt_df, supp_plt_df, med_yield))
  }

prep_ann_abund_for_plot <- function(df_exp_spec_noise_rec_fs,
                                    frac_reserves = c(0, 0.25, 0.50),
                                    f_fmsys = c(0.52, 1.00, 2.09)) {

  if (length(f_fmsys) < 1 | is.null(f_fmsys)) {
    stop("Please specify f_fmsy values to use")
  }

  if (length(frac_reserves) < 1 | is.null(frac_reserves)) {
    stop("Please specify frac_reserves values to use")
  }

  # make a sub function
    med_biomass <- df_exp_spec_noise_rec_fs %>%
      dplyr::filter(near(f_fmsy, 1), near(reserve_frac, 0))  %>%
      dplyr::group_by(
        experiment,
        noise,
        species,
        bh_alpha,
        rec_sd,
        reserve_frac,
        flep_ratio,
        sim_num,
        year
      ) %>%
      dplyr::summarise(ann_biomass = sum(biomass_at_age)) %>%
      dplyr::ungroup() %>%
      dplyr::summarise(med_biomass_f_fmsy = median(ann_biomass)) %>%
      .$med_biomass_f_fmsy

    # make a sub function
    main_plt_df <- df_exp_spec_noise_rec_fs %>%
      dplyr::filter(reduce(map(
        frac_reserves, near, x = reserve_frac, tol = 1e-4
      ), `|`)) %>%
      dplyr::group_by(reserve_frac, flep_ratio, f_fmsy, year) %>%
      dplyr::summarize(annual_biomass = sum(biomass_at_age)) %>%
      dplyr::ungroup() %>%
      dplyr::filter(reduce(map(
        f_fmsys, near, x = f_fmsy, tol = 1e-4
      ), `|`))

    # make a sub function
    supp_plt_df <- main_plt_df %>%
      dplyr::mutate(below_thresh = if_else(annual_biomass <= med_biomass, 1, 0)) %>%
      dplyr::group_by(reserve_frac, f_fmsy) %>%
      dplyr::summarize(
        mean_biomass = round(mean(annual_biomass, na.rm = TRUE), 2),
        mean_biomass_lab = paste0("Mean biomass: ", mean_biomass),
        sd_biomass = round(sd(annual_biomass, na.rm = TRUE), 2),
        sd_biomass_lab = paste0("SD biomass: ", sd_biomass),
        cv_biomass = round(sd(annual_biomass, na.rm = TRUE)/mean(annual_biomass, na.rm = TRUE), 2),
        cv_biomass_lab = paste0("CV yield: ", cv_biomass),
        yrs_below_thresh = paste("YR below thresh: ", sum(below_thresh))
      )
    return(list(main_plt_df, supp_plt_df, med_biomass))
  }

facet_yield_ts_plots <- function(main_plt_df,
                                 sub_plt_df,
                                 yield_thresh,
                                 text = TRUE,
                                 exp_deets_df,
                                 x_min = 600,
                                 x_max = 635,
                                 x_text_val = (x_max) - (x_max - x_min)/2) {

  plt_species <- unique(exp_deets_df$species)
  plt_dispersal <- unique(exp_deets_df$experiment)
  plt_noise <- unique(exp_deets_df$noise)
  plt_alpha <- round(unique(exp_deets_df$bh_alpha), 3)
  plt_sd <- unique(exp_deets_df$rec_sd)

  title_string <- paste0("Simulation results for: ", plt_species, 
                         " with ", plt_dispersal, " and ", plt_noise, " noise. \n", 
                         # "Pop dy parms: \n- BH alpha = ", plt_alpha, 
                         "\n- noise sd = ", plt_sd)

  quants_df <- main_plt_df %>%
    dplyr::group_by(reserve_frac, f_fmsy) %>%
    summarize(quant_80 = quantile(annual_yield, 0.8), 
              quant_20 = quantile(annual_yield, 0.2))

  ggplot(main_plt_df, aes(x = year, y = annual_yield)) +
    facet_grid(reserve_frac ~ f_fmsy,
               labeller = labeller(f_fmsy = label_both,
                                   reserve_frac = label_both)) +
    geom_ribbon(aes(
      ymin = yield_thresh,
      ymax = pmin(annual_yield, yield_thresh),
      fill = "Yield lower"
    )) +
    geom_ribbon(aes(
      ymin = annual_yield,
      ymax = pmin(annual_yield, yield_thresh),
      fill = "Yield higher"
    )) +
    scale_fill_manual(values = c("black", "white")) +
    {if (text) geom_text(
      data = sub_plt_df,
      x = x_text_val,
      y = 9500,
      size = 2.5,
      aes(label = mean_yield_lab)
    )} +
    {if (text) geom_text(
      data = sub_plt_df,
      x = x_text_val,
      y = 11500,
      size = 2.5,
      aes(label = sd_yield_lab)
    )} +
    {if (text) geom_text(
      data = sub_plt_df,
      x = x_text_val,
      y = 12500,
      size = 2.5,
      aes(label = yrs_below_thresh)
    )} +
    {if (text) geom_text(
      data = sub_plt_df,
      x = x_text_val,
      y = 10500,
      size = 2.5,
      aes(label = cv_yield_lab)
    )} +
    labs(fill = "Above/Below Yield threshold") +
    ggtitle(title_string) +
    xlab("Year") +
    ylab("Yield (metric tons)") +
    # ylim(0, 15000) +
    ylim(0, 12500) +
    xlim(x_min, x_max) +
    labs(colour = "Reserve\nfraction") +
    theme_grey() +
    geom_hline(data = quants_df, aes(yintercept = quant_80), colour = "orange", linewidth = 1.1, linetype = 1) +
    geom_hline(data = quants_df, aes(yintercept = quant_20), colour = "orange", linewidth = 1.1, linetype = 1) +
    theme(
      axis.line = element_line(colour = "black"),
      panel.background = element_rect(fill = "grey60"),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      panel.border = element_blank(),
      legend.position="none"
    )
}

facet_biomass_ys_plots <- function(main_plt_df,
                                 sub_plt_df,
                                 biomass_thresh,
                                 exp_deets_df,
                                 x_text_val = 550) {

  plt_species <- unique(exp_deets_df$species)
  plt_dispersal <- unique(exp_deets_df$experiment)
  plt_noise <- unique(exp_deets_df$noise)
  plt_alpha <- round(unique(exp_deets_df$bh_alpha), 3)
  plt_sd <- unique(exp_deets_df$rec_sd)

  title_string <- paste0("Simulation results for: ", plt_species, 
                         " with ", plt_dispersal, " and ", plt_noise, " noise. \n", 
                         # "Pop dy parms: \n- BH alpha = ", plt_alpha, 
                         "\n- noise sd = ", plt_sd)

  quants_df <- main_plt_df %>%
    dplyr::group_by(reserve_frac, f_fmsy) %>%
    summarize(quant_80 = quantile(annual_biomass, 0.8),
              quant_20 = quantile(annual_biomass, 0.2))

  ggplot(main_plt_df, aes(x = year, y = annual_biomass)) +
    facet_grid(reserve_frac ~ f_fmsy,
               labeller = labeller(f_fmsy = label_both,
                                   reserve_frac = label_both)) +
    geom_ribbon(aes(
      ymin = biomass_thresh,
      ymax = pmin(annual_biomass, biomass_thresh),
      fill = "Yield lower"
    )) +
    geom_ribbon(aes(
      ymin = annual_biomass,
      ymax = pmin(annual_biomass, biomass_thresh),
      fill = "Yield higher"
    )) +
    scale_fill_manual(values = c("black", "white")) +
    # geom_text(
    #   data = sub_plt_df,
    #   x = x_text_val,
    #   y = 11000,
    #   aes(label = mean_biomass_lab)
    # ) +
    # geom_text(
    #   data = sub_plt_df,
    #   x = x_text_val,
    #   y = 12500,
    #   aes(label = sd_biomass_lab)
    # ) +
    # geom_text(
    #   data = sub_plt_df,
    #   x = x_text_val,
    #   y = 14000,
    #   aes(label = yrs_below_thresh)
    # ) +
    # geom_text(
    #   data = sub_plt_df,
    #   x = x_text_val,
    #   y = 9500,
    #   aes(label = cv_biomass_lab)
    # ) +
    labs(fill = "Above/Below Biomass threshold") +
    ggtitle(title_string) +
    xlab("Year") +
    ylab("Biomass (metric tons)") +
    ylim(0, NA) +
    labs(colour = "Reserve\nfraction") +
    theme_grey() +
    geom_hline(data = quants_df, aes(yintercept = quant_80), colour = "orange", linewidth = 1.1, linetype = 1) +
    geom_hline(data = quants_df, aes(yintercept = quant_20), colour = "orange", linewidth = 1.1, linetype = 1) +
    theme(
      axis.line = element_line(colour = "black"),
      panel.background = element_rect(fill = "grey60"),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      panel.border = element_blank()
    )
}

# Recruits TS above/below thresh time series

# Recruit time series
prep_ann_recs_for_plot <- function(df_exp_spec_noise_rec_fs,
                                   fleps_2_fs,
                                   f_maxes_yield,
                                   frac_reserves = c(0, 0.25, 0.50),
                                   f_fmsys = c(0, 0.52, 1.00, 1,43, 2.09),
                                   sp_parms,
                                   spc) {
  if (length(f_fmsys) < 1 | is.null(f_fmsys)) {
    stop("Please specify f_fmsy values to use")
  }

  if (length(frac_reserves) < 1 | is.null(frac_reserves)) {
    stop("Please specify frac_reserves values to use")
  }

  t_c <- sp_parms[which(sp_parms$species == spc), ]$A_fish

  main_plt_df <- df_exp_spec_noise_rec_fs |>
    left_join(., y = fleps_2_fs, by = join_by(flep_ratio == flep_vals)) |>
    dplyr::mutate(f_fmsy = round(f_vals / f_maxes_yield, 2)) |>
    dplyr::filter(reduce(map(
      frac_reserves, near, x = reserve_frac, tol = 1e-4
    ), `|`),
    reduce(map(
      f_fmsys, near, x = f_fmsy, tol = 1e-4
    ), `|`))

  # make a sub function
  recruit_thresh <- main_plt_df %>%
    dplyr::filter(near(f_fmsy, 1), near(reserve_frac, 0)) %>%
    dplyr::summarize(med_recs_f_fmsy = median(recruits)) %>%
    .$med_recs_f_fmsy

  supp_plt_df <- main_plt_df %>%
  dplyr::filter(year > 500) %>%
  dplyr::mutate(below_thresh = if_else(recruits <= recruit_thresh, 1, 0)) %>%
  dplyr::group_by(reserve_frac, f_fmsy) |>
      dplyr::summarize(
        mean_recruits = round(mean(recruits, na.rm = TRUE), 2),
        sd_recruits = round(sd(recruits, na.rm = TRUE), 2),
        cv_recruits = round(sd_recruits/mean_recruits, 2),
        yrs_below_thresh = paste("YR below thresh: ", sum(below_thresh))
      ) |>
        dplyr::ungroup() |>
        dplyr::mutate(mean_recruits_lab = paste0("Mean recruits: ", format(mean_recruits, 
                                                             big.mark = ",",
                                                             scientific = FALSE)),
                      sd_recruits_lab = paste0("SD recruits: ", format(sd_recruits, 
                                                             big.mark = ",",
                                                             scientific = FALSE)),
                      cv_recruits_lab = paste0("CV recruits: ", format(cv_recruits, 
                                                             big.mark = ",",
                                                             scientific = FALSE)))
  return(list(main_plt_df, supp_plt_df, recruit_thresh))
}

facet_recs_ts_plots <- function(main_plt_df,
                                sub_plt_df,
                                yield_sub_plt,
                                recruit_thresh,
                                exp_deets_df,
                                x_text_val = 550) {
  ggplot(data = main_plt_df |> filter(year > 500),
       aes(x = year, y = recruits)) +
  geom_ribbon(aes(
      ymin = recruit_thresh,
      ymax = pmin(recruits, recruit_thresh),
      fill = "Recruits lower"
    )) +
    geom_ribbon(aes(
      ymin = recruits,
      ymax = pmin(recruits, recruit_thresh),
      fill = "Recruits higher"
    )) +
    scale_fill_manual(values = c("black", "white")) +
  facet_grid(reserve_frac ~ f_fmsy,
             labeller = labeller(f_fmsy = label_both,
                                 reserve_frac = label_both)) +
      geom_text(
      data = sub_plt_df,
      x = x_text_val,
      y = 4E6,
      aes(label = mean_recruits_lab),
      size = 3
    ) +
    geom_text(
      data = sub_plt_df,
      x = x_text_val,
      y = 4.5E6,
      aes(label = sd_recruits_lab),
      size = 3
    ) +
    geom_text(
      data = sub_plt_df,
      x = x_text_val,
      y = 5E6,
      aes(label = yrs_below_thresh),
      size = 3
    ) +
    geom_text(
      data = sub_plt_df,
      x = x_text_val,
      y = 3.5E6,
      aes(label = cv_recruits_lab),
      size = 3
    ) + 
    geom_text(
      data = yield_sub_plt,
      x = x_text_val,
      y = 5.5E6,
      aes(label = mean_yield_lab),
      size = 3
    ) +
    labs(fill = "Above/Below Recruit threshold") +
    ylim(0, 7e6) +
  theme_grey()  +
  theme(
    axis.line = element_line(colour = "black"),
    panel.background = element_rect(fill = "grey60"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank()
  ) +
  ylab("Recruits") +
  xlab("Year")

}

# Contour plot functions: yield

get_sp_disp_noise_yield_gcols <- function(sp_disp_noise_yield_prep) {
  sp_disp_noise_yield_prep_gcols <- sp_disp_noise_yield_prep %>%
    select(all_of(
      c(
        "reserve_frac",
        "flep_ratio",
        "f_vals",
        "f_fmsy",
        "year",
        "age",
        "yield"
      )
    ))

  return(sp_disp_noise_yield_prep_gcols)
}

get_sp_disp_noise_abund_gcols <- function(sp_disp_noise_abund_prep) {
  sp_disp_noise_abund_prep_gcols <- sp_disp_noise_abund_prep %>%
    select(all_of(
      c(
        "reserve_frac",
        "flep_ratio",
        "f_vals",
        "f_fmsy",
        "year",
        "age",
        "biomass_at_age"
      )
    ))

  return(sp_disp_noise_abund_prep_gcols)
}

calc_sp_disp_noise_gcols_ann_yield <- function(sp_disp_noise_yield_prep_gcols) {
  sp_disp_noise_yield_gcols_ann_yield <- sp_disp_noise_yield_prep_gcols %>%
  group_by(reserve_frac, f_vals, f_fmsy, year) %>%
  summarise(ann_yield = sum(yield)) %>%
  ungroup() 
  return(sp_disp_noise_yield_gcols_ann_yield)
}

calc_sp_disp_noise_gcols_ann_abund <- function(sp_disp_noise_abund_prep_gcols) {
  sp_disp_noise_abund_gcols_ann<- sp_disp_noise_abund_prep_gcols %>%
  group_by(reserve_frac, f_vals, f_fmsy, year) %>%
  summarise(ann_biomass = sum(biomass_at_age)) %>%
  ungroup() 
  return(sp_disp_noise_abund_gcols_ann)
}

calc_sp_disp_noise_yield_gcols_med_ann_yield <- function(sp_disp_noise_yield_prep_gcols) {
  sp_disp_noise_yield_gcols_med_ann_yield <- sp_disp_noise_yield_prep_gcols %>%
  group_by(reserve_frac, f_vals, f_fmsy, year) %>%
  summarise(ann_yield = sum(yield)) %>%
  ungroup() %>%
  group_by(reserve_frac, f_vals,  f_fmsy) %>%
  summarise(median_ann_yield = median(ann_yield)) %>%
  ungroup()
  return(sp_disp_noise_yield_gcols_med_ann_yield)
}

calc_sp_disp_noise_abund_gcols_med_ann_abund <- function(sp_disp_noise_abund_prep_gcols) {
  sp_disp_noise_abund_gcols_med_ann_abund <- sp_disp_noise_abund_prep_gcols %>%
  group_by(reserve_frac, f_vals, f_fmsy, year) %>%
  summarise(ann_biomass = sum(biomass_at_age)) %>%
  ungroup() %>%
  group_by(reserve_frac, f_vals,  f_fmsy) %>%
  summarise(median_ann_biomass = median(ann_biomass)) %>%
  ungroup()
  return(sp_disp_noise_abund_gcols_med_ann_abund)
}

get_f_maxes_med_ann_yield <- function(sp_disp_noise_yield_gcols_med_ann_yield) {
  f_max_median_ann_yield <- sp_disp_noise_yield_gcols_med_ann_yield %>%
  filter(median_ann_yield == max(median_ann_yield)) %>%
  .$f_vals

  try(if(length(f_max_median_ann_yield) == 0) stop("no F that maxes yield was found"))

  if (length(f_max_median_ann_yield) > 1) {

    return(f_max_median_ann_yield[1]) 
  }

  return(f_max_median_ann_yield)
}

calc_sp_disp_noise_yield_gcols_ann_yield_ffmsy <- function(sp_disp_noise_yield_gcols_ann_yield, 
                                                      f_max_median_ann_yield) {

  sp_disp_noise_yield_gcols_ann_yield_ffmsy <- sp_disp_noise_yield_gcols_ann_yield %>%
    mutate(f_fmsy = f_vals / f_max_median_ann_yield)

  return(sp_disp_noise_yield_gcols_ann_yield_ffmsy)
}

calc_sp_disp_noise_abund_gcols_ann_abund_ffmsy <- function(sp_disp_noise_abund_gcols_ann_abund, 
                                                      f_max_median_ann_yield) {

  sp_disp_noise_abund_gcols_ann_abund_ffmsy <- sp_disp_noise_abund_gcols_ann_abund %>%
    mutate(f_fmsy = f_vals / f_max_median_ann_yield)

  return(sp_disp_noise_abund_gcols_ann_abund_ffmsy)
}

calc_sp_disp_noise_yield_frac_below_med_max_yield <- function(sp_disp_noise_yield_gcols_ann_yield_ffmsy,
                                                              max_median_ann_yield) {

  # if (is.na(max_median_ann_yield)) {
  #   max_median_ann_yield <- sp_disp_noise_yield_gcols_ann_yield_ffmsy %>% 
  #   group_by(reserve_frac, f_vals, f_fmsy) %>%
  #   summarise(median_ann_yield = median(ann_yield)) %>%
  #   ungroup() %>%
  #   filter(median_ann_yield == max(median_ann_yield)) %>% 
  #   .$median_ann_yield
  # }

  sp_disp_noise_yield_frac_below_med_max_yield <-
      sp_disp_noise_yield_gcols_ann_yield_ffmsy %>%
      mutate(below_yield_thresh = if_else(ann_yield < max_median_ann_yield, 1, 0)) %>%
      group_by(reserve_frac, f_fmsy) %>%
      summarise(n_yrs_below = sum(below_yield_thresh)) %>%
      ungroup() %>%
      mutate(frac_below_med_yield_f_max = round(n_yrs_below / sim_len, 3))
    return(sp_disp_noise_yield_frac_below_med_max_yield)
}

calc_sp_disp_noise_abund_frac_below_med_max_abund <- function(sp_disp_noise_abund_gcols_ann_abund_ffmsy,
                                                              max_median_ann_abund) {

  # if (is.na(max_median_ann_abund)) {
  #   max_median_ann_abund <- sp_disp_noise_abund_gcols_ann_abund_ffmsy %>% 
  #   group_by(reserve_frac, f_vals, f_fmsy) %>%
  #   summarise(median_ann_biomass = median(ann_biomass)) %>%
  #   ungroup() %>%
  #   filter(median_ann_biomass == max(median_ann_biomass)) %>% 
  #   .$median_ann_biomass
  # }

  sp_disp_noise_abund_frac_below_med_max_abund <-
      sp_disp_noise_abund_gcols_ann_abund_ffmsy %>%
      mutate(below_thresh_ann_yield = if_else(ann_biomass < max_median_ann_abund, 1, 0)) %>%
      group_by(reserve_frac, f_fmsy) %>%
      summarise(n_yrs_below = sum(below_thresh_ann_yield)) %>%
      ungroup() %>%
      mutate(frac_below_med_abund_f_max = round(n_yrs_below / sim_len, 3))
    return(sp_disp_noise_abund_frac_below_med_max_abund)
}

make_contour_plt_sp_disp_noise_frac_below_med_max_yield <-
  function(sp_disp_noise_yield_frac_below_med_max_yield,
           species_name,
           dispersal,
           noise_type,
           max_f_fmsy,
           max_reserve_frac) {
    sp_disp_noise_yield_frac_below_med_max_yield_plt <-
      sp_disp_noise_yield_frac_below_med_max_yield %>%
      filter(reserve_frac <= max_reserve_frac, f_fmsy <= max_f_fmsy) %>%
      ggplot(data = .,
             aes(x = reserve_frac, y = f_fmsy, z = frac_below_med_yield_f_max)) +
      geom_contour_filled(bins = 12) +
      ggtitle(paste0(
        species_name,
        " ",
        dispersal,
        " ",
        noise_type,
        ":\nTime below median max yield"
      )) +
      xlab("Reserve fraction") +
      ylab("F/Fmsy") +
      scale_fill_viridis_d(option = "D", direction = -1) +
      geom_segment(aes(
        x = 0,
        xend = max_reserve_frac,
        y = 1,
        yend = 1
      ), colour = "black",
      linetype = 2) +
      scale_x_continuous(expand = expansion(mult = 0, add = 0)) +
      scale_y_continuous(expand = expansion(mult = 0, add = 0)) +
      theme_bw()

    return(sp_disp_noise_yield_frac_below_med_max_yield_plt)
  }

make_contour_plt_sp_disp_noise_frac_below_med_max_abund <-
  function(sp_disp_noise_frac_below_med_max_abund,
           species_name,
           dispersal,
           noise_type,
           max_f_fmsy,
           max_reserve_frac) {
    sp_disp_noise_frac_below_med_max_abund_plt <-
      sp_disp_noise_frac_below_med_max_abund %>%
      filter(reserve_frac <= max_reserve_frac, f_fmsy <= max_f_fmsy) %>%
      ggplot(data = .,
             aes(x = reserve_frac, y = f_fmsy, z = frac_below_med_abund_f_max)) +
      geom_contour_filled(bins = 12) +
      ggtitle(
        paste0(
          species_name,
          " ",
          dispersal,
          " ",
          noise_type,
          ":\nTime below median max biomass"
        )
      ) +
      xlab("Reserve fraction") +
      ylab("F/Fmsy") +
      scale_fill_viridis_d(option = "D", direction = -1) +
      geom_segment(
        aes(
          x = 0,
          xend = max_reserve_frac,
          y = 1,
          yend = 1
        ),
        colour = "black",
        linetype = 2
      ) +
      scale_x_continuous(expand = expansion(mult = 0, add = 0)) +
      scale_y_continuous(expand = expansion(mult = 0, add = 0)) +
      theme_bw()

    return(sp_disp_noise_frac_below_med_max_abund_plt)
}

# Contour plot functions: recruits

calc_sp_disp_noise_recs_ffmsy <- function(sp_disp_noise_recs, 
                                          fleps_2_fs,
                                          f_maxes_med_ann_yield) {
  sp_disp_noise_recs_ffmsy <- sp_disp_noise_recs %>%
    mutate(flep_ratio_ch = as.character(flep_ratio)) %>%
    left_join(., y = fleps_2_fs, by = join_by(flep_ratio_ch == flep_vals_ch)) %>%
    dplyr::mutate(f_fmsy = round(f_vals / f_maxes_med_ann_yield, 2))
  return(sp_disp_noise_recs_ffmsy)
}

calc_recruit_thresh <- function(sp_disp_noise_recs_ffmsy) {
  recruit_thresh <- sp_disp_noise_recs_ffmsy %>%
  dplyr::filter(near(f_fmsy, 1), near(reserve_frac, 0)) %>%
  dplyr::summarize(med_recs_f_fmsy = median(recruits)) %>%
  .$med_recs_f_fmsy
  return(recruit_thresh)
}

calc_time_below_rec_thresh <- function(sp_disp_noise_recs_ffmsy, 
                                       recruit_thresh) {
  sp_disp_noise_recs_ffmsy %>%
  arrange(reserve_frac, f_fmsy, year) %>%
  mutate(below_threshold = recruits <= recruit_thresh) %>%
  group_by(reserve_frac, f_fmsy) %>%
  summarize(num_below_threshold = sum(below_threshold), 
            frac_below_threshold = num_below_threshold/sim_len)
}

make_contour_sp_disp_noise_frac_below_rec_thresh <-
  function(sp_disp_noise_recs_below_thresh,
           species_name,
           dispersal,
           noise_type,
           max_f_fmsy = 2.5 ,
           max_reserve_frac = 0.5,
           n_bins = 12) {

    recs_below_threshold_plt <- sp_disp_noise_recs_below_thresh %>%
      filter(reserve_frac <= max_reserve_frac, f_fmsy <= max_f_fmsy) %>%
      ggplot(data = .,
             aes(x = reserve_frac, y = f_fmsy, z = frac_below_threshold)) +
      geom_contour_filled(bins = n_bins) +
      ggtitle(paste0(
        species_name,
        " ",
        dispersal,
        " ",
        noise_type,
        ":\nTime below threshold"
      )) +
      xlab("Reserve fraction") +
      ylab("F/Fmsy") +
      scale_fill_viridis_d(option = "D", direction = -1) +
      geom_segment(aes(
        x = 0,
        xend = max_reserve_frac,
        y = 1,
        yend = 1
      ),
      colour = "black",
      linetype = 2) +
      scale_x_continuous(expand = expansion(mult = 0, add = 0)) +
      scale_y_continuous(expand = expansion(mult = 0, add = 0))

    return(recs_below_threshold_plt)

  }

```

```{r}
#| label: recs-abund-yield
#| echo: FALSE

recs_spec_disp_noise <- qry_results_table_agg(
  ddb_table = exp, #"larval_pool_blue_rockfish_mei_135_75",
  metric = "recruits",
  burn_in = burn_in_parm,
  sim_len = sim_len_parm,
  reserve_fracs = NULL,
  flep_ratios = NULL) %>%
  mutate(flep_ratio_ch = as.character(flep_ratio))

waa_df <- tibble(age = 1:length(waa_g),
                 "waa" = waa_g)

abund_spec_disp_noise <- qry_results_table_agg(
  ddb_table = exp, #"larval_pool_blue_rockfish_mei_135_75",
  metric = "abundance",
  burn_in = burn_in_parm,
  sim_len = sim_len_parm,
  reserve_fracs = NULL,
  flep_ratios = NULL) %>%
  tibble() %>%
  left_join(x = ., y = waa_df, by = "age") %>%
  mutate(biomass_at_age = (abundance * waa) / 1e6,
         flep_ratio_ch = as.character(flep_ratio)) %>%
  select(-waa, -abundance)

yield_spec_disp_noise <- qry_results_table_agg(
  ddb_table = exp, #"larval_pool_blue_rockfish_mei_135_75",
  metric = "yield",
  burn_in = burn_in_parm,
  sim_len = sim_len_parm,
  reserve_fracs = NULL,
  flep_ratios = NULL) %>% 
  mutate(yield = yield / 1e6, 
         flep_ratio_ch = as.character(flep_ratio))

```

## Recruits

#### Time series plots

```{r}
#| label: recruit-time-series-plots
#| echo: FALSE
#| fig-width: 15
#| fig-height: 8
#| fig-column: screen

F_max_yield <-
  get_f_maxes_yield(yield_disp_sp_noise = yield_spec_disp_noise,
                    fleps_2_fs = spec_fleps_2_fs)

# yield_brf_lp_white_135_75_prep

yield_spec_disp_noise_prep <- make_df_yield_exp_spec_noise_rec_fs(
  df_exp_spec_noise_rec = yield_spec_disp_noise,
  fleps_2_fs = spec_fleps_2_fs,
  f_maxes_yield = F_max_yield
)

yield_sub_plt_df <- prep_ann_yield_for_plot(
  df_exp_spec_noise_rec = yield_spec_disp_noise_prep,
  frac_reserves = c(0, 0.25, 0.50),
  f_fmsys = c(0, 0.52, 1.00, 1.43, 2.09)
)[[2]]

recs_spec_disp_noise_plt <-
  prep_ann_recs_for_plot(
    df_exp_spec_noise_rec_fs = recs_spec_disp_noise,
    fleps_2_fs = spec_fleps_2_fs,
    f_maxes_yield = F_max_yield,
    frac_reserves = c(0, 0.25, 0.50),
    f_fmsys = c(0, 0.52, 1.00, 1.43, 2.09),
    sp_parms = species_parms,
    spc = nice_species
  )

spec_disp_noise_135_075_alpha_sd <-
  get_exp_details(df_after_qry_results_table_agg = recs_spec_disp_noise)

recs_ts_plt <- facet_recs_ts_plots(
  main_plt_df = recs_spec_disp_noise_plt[[1]],
  sub_plt_df = recs_spec_disp_noise_plt[[2]],
  yield_sub_plt = yield_sub_plt_df,
  recruit_thresh = recs_spec_disp_noise_plt[[3]],
  exp_deets_df = spec_disp_noise_135_075_alpha_sd,
  x_text_val = 550
)

print(recs_ts_plt)

postscript(file = file.path(out_path, paste0(exp, "-recs-ts-plt.eps")), 
           width = 8,
           heigh = 6)
print(recs_ts_plt)
invisible(dev.off())

png(
  file = file.path(out_path, paste0(exp, "-recs-ts-plt.png")),
  width = 8,
  heigh = 6,
  units = "in", res = 600
)
print(recs_ts_plt)
invisible(dev.off())

```

#### Contour plots

```{r}
#| label: recruit-contour-plots
#| echo: FALSE
#| fig-width: 15
#| fig-height: 8
#| fig-column: screen

# yield_spec_disp_noise_prep
yield_spec_disp_noise_prep_gcols <- get_sp_disp_noise_yield_gcols(sp_disp_noise_yield_prep = yield_spec_disp_noise_prep)

yield_spec_disp_noise_gcols_ann <- 
calc_sp_disp_noise_gcols_ann_yield(sp_disp_noise_yield_prep_gcols = yield_spec_disp_noise_prep_gcols)

yield_spec_disp_noise_gcols_med_ann <- calc_sp_disp_noise_yield_gcols_med_ann_yield(sp_disp_noise_yield_prep_gcols = yield_spec_disp_noise_prep_gcols)

# spec_disp_noise_f_maxes_med_ann_yield <- get_f_maxes_med_ann_yield(sp_disp_noise_yield_gcols_med_ann_yield = yield_spec_disp_noise_gcols_med_ann)

recs_spec_disp_noise_ffmsy <-
  calc_sp_disp_noise_recs_ffmsy(sp_disp_noise_recs = recs_spec_disp_noise,
                                fleps_2_fs = spec_fleps_2_fs,
                                f_maxes_med_ann_yield = F_max_yield)

recs_spec_disp_rec_thresh <- calc_recruit_thresh(sp_disp_noise_recs_ffmsy = recs_spec_disp_noise_ffmsy)

recs_spec_disp_noise_below_thresh <-
  calc_time_below_rec_thresh(sp_disp_noise_recs_ffmsy = recs_spec_disp_noise_ffmsy,
                             recruit_thresh = recs_spec_disp_rec_thresh)

rec_contour_plt <- make_contour_sp_disp_noise_frac_below_rec_thresh(
  sp_disp_noise_recs_below_thresh = recs_spec_disp_noise_below_thresh,
  species_name = nice_species,
  dispersal = nice_dispersal,
  noise_type = nice_noise,
  max_f_fmsy = 2.5,
  max_reserve_frac = 0.5,
  n_bins = 12
)

print(rec_contour_plt)

postscript(file = file.path(out_path, paste0(exp, "-recs-contour-plt.eps")), 
           width = 8,
           heigh = 6)
print(rec_contour_plt)
invisible(dev.off())

png(
  file = file.path(out_path, paste0(exp, "-recs-contour-plt.png")),
  width = 8,
  heigh = 6,
  units = "in", res = 600
)
print(rec_contour_plt)
invisible(dev.off())

```

## Abundance

#### Time series plots

```{r}
#| label: abundance-time-series-plots
#| echo: FALSE
#| fig-width: 15
#| fig-height: 8
#| fig-column: screen
# Needs
# 1. abund_spec_disp_noise
# 2. (brf_)F_max_yield(_75sd)

# abund_spec_disp_noise
# F_max_yield

abund_spec_disp_noise_alpha_sd <- get_exp_details(df_after_qry_results_table_agg= abund_spec_disp_noise)

abund_spec_disp_noise_prep <- make_df_exp_spec_noise_rec_fs(
  df_exp_spec_noise_rec = abund_spec_disp_noise,
  fleps_2_fs = spec_fleps_2_fs,
  f_maxes_yield = F_max_yield
)

abund_spec_disp_noise_plt <-
  prep_ann_abund_for_plot(
    df_exp_spec_noise_rec = abund_spec_disp_noise_prep,
    frac_reserves = c(0, 0.25, 0.50),
    f_fmsys = c(0.52, 1.00, 1.43, 2.09)
  )

abund_ts_plt <- facet_biomass_ys_plots(main_plt_df = abund_spec_disp_noise_plt[[1]],
                     sub_plt_df = abund_spec_disp_noise_plt[[2]],
                     biomass_thresh = abund_spec_disp_noise_plt[[3]],
                     exp_deets_df = abund_spec_disp_noise_alpha_sd,
                     x_text_val = x_text_position_parm)

print(abund_ts_plt)

postscript(file = file.path(out_path, paste0(exp, "-abund-ts-plt.eps")), 
           width = 8,
           heigh = 6)
print(abund_ts_plt)
invisible(dev.off())

png(
  file = file.path(out_path, paste0(exp, "-abund-ts-plt.png")),
  width = 8,
  heigh = 6,
  units = "in", res = 600
)
print(abund_ts_plt)
invisible(dev.off())

```

#### Contour plots

```{r}
#| label: abundance-contour-plots
#| echo: FALSE
#| fig-width: 15
#| fig-height: 8
#| fig-column: screen

abund_spec_disp_noise_prep_gcols <- get_sp_disp_noise_abund_gcols(sp_disp_noise_abund_prep = abund_spec_disp_noise_prep)

abund_spec_disp_noise_prep_gcols_ann <- 
calc_sp_disp_noise_gcols_ann_abund(sp_disp_noise_abund_prep_gcols = abund_spec_disp_noise_prep_gcols)

abund_spec_disp_noise_prep_gcols_med_ann <- calc_sp_disp_noise_abund_gcols_med_ann_abund(sp_disp_noise_abund_prep_gcols = abund_spec_disp_noise_prep_gcols)

spec_disp_noise_ann_abund_ffmsy <- calc_sp_disp_noise_abund_gcols_ann_abund_ffmsy(
  sp_disp_noise_abund_gcols_ann_abund = abund_spec_disp_noise_prep_gcols_med_ann,
  f_max_median_ann_yield = F_max_yield
)

max_median_ann_abund_value <- spec_disp_noise_ann_abund_ffmsy %>% filter(near(f_fmsy, 1)) %>% .$median_ann_biomass

spec_disp_noise_prep_frac_below_med_max_abund <- calc_sp_disp_noise_abund_frac_below_med_max_abund(sp_disp_noise_abund_gcols_ann_abund_ffmsy = abund_spec_disp_noise_prep_gcols_ann, max_median_ann_abund = max_median_ann_abund_value)

spec_disp_noise_prep_abund_contour_plt <- make_contour_plt_sp_disp_noise_frac_below_med_max_abund(
  sp_disp_noise_frac_below_med_max_abund = spec_disp_noise_prep_frac_below_med_max_abund,
  species_name = nice_species,
  dispersal = nice_dispersal,
  noise_type = nice_noise,
  max_f_fmsy = 2.5,
  max_reserve_frac = 0.5
)

print(spec_disp_noise_prep_abund_contour_plt)

postscript(file = file.path(out_path, paste0(exp, "-abund-contour-plt.eps")), 
           width = 8,
           heigh = 6)
print(spec_disp_noise_prep_abund_contour_plt)
invisible(dev.off())

png(
  file = file.path(out_path, paste0(exp, "-abund-contour-plt.png")),
  width = 8,
  heigh = 6,
  units = "in", res = 600
)
print(spec_disp_noise_prep_abund_contour_plt)
invisible(dev.off())

```

## Yield

#### Time series

```{r}
#| label: yield-time-series-plots
#| echo: FALSE
#| fig-width: 15
#| fig-height: 8
#| fig-column: screen

# Data input needs:
# 1. yield_spec_disp_noise
# 2. F_max_yield

# F_max_yield <-
#   get_f_maxes_yield(yield_disp_sp_noise = yield_spec_disp_noise,
#                     fleps_2_fs = spec_fleps_2_fs)

spec_disp_noise_alpha_sd <- get_exp_details(df_after_qry_results_table_agg= yield_spec_disp_noise)

yield_spec_disp_noise_prep <- make_df_exp_spec_noise_rec_fs(
  df_exp_spec_noise_rec = yield_spec_disp_noise,
  fleps_2_fs = spec_fleps_2_fs,
  f_maxes_yield = F_max_yield
)

yield_spec_disp_noise_plt <-
  prep_ann_yield_for_plot(
    df_exp_spec_noise_rec = yield_spec_disp_noise_prep,
    frac_reserves = c(0, 0.25, 0.50),
    f_fmsys = c(0.52, 1.00, 1.43, 2.09)
  )

yield_ts_plots <- facet_yield_ts_plots(main_plt_df = yield_spec_disp_noise_plt[[1]],
                     sub_plt_df = yield_spec_disp_noise_plt[[2]],
                     yield_thresh = yield_spec_disp_noise_plt[[3]],
                     exp_deets_df = spec_disp_noise_alpha_sd,
                     x_min = 600,
                     x_max = 635)

print(yield_ts_plots)

postscript(file = file.path(out_path, paste0(exp, "-yield-ts-plt.eps")), 
           width = 8,
           heigh = 6)
print(yield_ts_plots)
invisible(dev.off())

png(
  file = file.path(out_path, paste0(exp, "-yield-ts-plt.png")),
  width = 8,
  heigh = 6,
  units = "in", res = 600
)
print(yield_ts_plots)
invisible(dev.off())

# No text version
yield_ts_plots <- facet_yield_ts_plots(main_plt_df = yield_spec_disp_noise_plt[[1]],
                     sub_plt_df = yield_spec_disp_noise_plt[[2]],
                     yield_thresh = yield_spec_disp_noise_plt[[3]],
                     exp_deets_df = spec_disp_noise_alpha_sd,
                     text = FALSE,
                     x_min = 600,
                     x_max = 635)

postscript(file = file.path(out_path, paste0(exp, "-yield-ts-plt-no-text.eps")), 
           width = 8,
           heigh = 6)
print(yield_ts_plots)
invisible(dev.off())

png(
  file = file.path(out_path, paste0(exp, "-yield-ts-plt-no-text.png")),
  width = 8,
  heigh = 6,
  units = "in", res = 600
)
print(yield_ts_plots)
invisible(dev.off())

```

#### Contour

```{r}
#| label: yield-contour-plots
#| echo: FALSE
#| fig-width: 15
#| fig-height: 8
#| fig-column: screen

yield_spec_disp_noise_prep_gcols <- get_sp_disp_noise_yield_gcols(sp_disp_noise_yield_prep = yield_spec_disp_noise_prep)

yield_spec_disp_noise_prep_gcols_ann <- 
calc_sp_disp_noise_gcols_ann_yield(sp_disp_noise_yield_prep_gcols = yield_spec_disp_noise_prep_gcols)

yield_spec_disp_noise_prep_gcols_med_ann <- calc_sp_disp_noise_yield_gcols_med_ann_yield(sp_disp_noise_yield_prep_gcols = yield_spec_disp_noise_prep_gcols)

spec_disp_noise_ann_yield_ffmsy <- calc_sp_disp_noise_yield_gcols_ann_yield_ffmsy(
  sp_disp_noise_yield_gcols_ann_yield = yield_spec_disp_noise_prep_gcols_med_ann,
  f_max_median_ann_yield = F_max_yield)

spec_disp_noise_prep_frac_below_med_max_yield <- calc_sp_disp_noise_yield_frac_below_med_max_yield(sp_disp_noise_yield_gcols_ann_yield_ffmsy = yield_spec_disp_noise_prep_gcols_ann, max_median_ann_yield = spec_disp_noise_ann_yield_ffmsy %>% filter(near(f_vals, F_max_yield, 0.001), near(reserve_frac, 0, 0.001)) %>% .$median_ann_yield)

spec_disp_noise_prep_yield_contour_plt <- make_contour_plt_sp_disp_noise_frac_below_med_max_yield(
  sp_disp_noise_yield_frac_below_med_max_yield = spec_disp_noise_prep_frac_below_med_max_yield,
  species_name = nice_species,
  dispersal = nice_dispersal,
  noise_type = nice_noise,
  max_f_fmsy = 2.5,
  max_reserve_frac = 0.5
)

print(spec_disp_noise_prep_yield_contour_plt)

postscript(file = file.path(out_path, paste0(exp, "-yield-contour-plt.eps")), 
           width = 8,
           heigh = 6)
print(spec_disp_noise_prep_yield_contour_plt)
invisible(dev.off())

png(
  file = file.path(out_path, paste0(exp, "-yield-contour-plt.png")),
  width = 8,
  heigh = 6,
  units = "in", res = 600
)
print(spec_disp_noise_prep_yield_contour_plt)
invisible(dev.off())

```

## Lagged-recruits ($t_c$) and yield

#### Time series

```{r}
#| label: lagged-recs-yield-ts-plots
#| echo: FALSE
#| fig-width: 15
#| fig-height: 8
#| fig-column: screen

# yield_spec_disp_noise_prep_gcols
# recs_spec_disp_noise_ffmsy

frac_reserves <- c(0, 0.25, 0.50)
f_fmsys <- c(0.52, 1.00, 2.09)

t_c <- species_parms[which(species_parms$species == nice_species),]$A_fish

recs_spec_disp_noise_ffmsy_plt <- recs_spec_disp_noise_ffmsy %>%
  dplyr::filter(reduce(map(
    frac_reserves, near, x = reserve_frac, tol = 1e-4
  ), `|`),
  reduce(map(
    f_fmsys, near, x = f_fmsy, tol = 1e-4
  ), `|`)) %>%
  mutate(lagged_year = year + t_c)

yield_spec_disp_noise_prep_plt <- yield_spec_disp_noise_prep_gcols %>%
  dplyr::group_by(reserve_frac, flep_ratio, f_fmsy, year) %>%
  dplyr::summarize(annual_yield = sum(yield)) %>%
  dplyr::ungroup() %>%
  dplyr::filter(reduce(map(
    frac_reserves, near, x = reserve_frac, tol = 1e-4
  ), `|`),
  reduce(map(
    f_fmsys, near, x = f_fmsy, tol = 1e-4
  ), `|`))

plt_start_yr <- 600
plt_end_yr <- 635

scaled_yield_plt <- yield_spec_disp_noise_prep_plt %>%
  group_by(reserve_frac, f_fmsy) %>%
  mutate(scaled_yield = scale(annual_yield)[,1]) %>%
  ungroup()

scaled_recs_plt <- recs_spec_disp_noise_ffmsy_plt %>% 
  group_by(reserve_frac, f_fmsy) %>%
  mutate(scaled_recruits = scale(recruits)[,1]) %>%
  ungroup()

scaled_yield_plt_4corr <- scaled_yield_plt %>%
  mutate(reserve_frac = as.character(reserve_frac), 
         f_fmsy = as.character(f_fmsy)) %>%
  select(reserve_frac, f_fmsy, year, scaled_yield)

scaled_recs_plt_4corr <- scaled_recs_plt %>%
  mutate(reserve_frac = as.character(reserve_frac), 
         f_fmsy = as.character(f_fmsy))  %>%
  select(reserve_frac, f_fmsy, lagged_year, scaled_recruits)

corr_vals <- scaled_yield_plt_4corr %>%
  left_join(scaled_recs_plt_4corr,
            by = c("reserve_frac" = "reserve_frac",
                   "f_fmsy" = "f_fmsy",
                   "year" = "lagged_year")) %>%
  filter(.,complete.cases(reserve_frac, f_fmsy, year, scaled_yield, scaled_recruits)) %>%
  group_by(reserve_frac, f_fmsy) %>%
  summarise(corr = cor(scaled_yield, scaled_recruits))

lagged_recs_yield_plt <- ggplot(scaled_yield_plt %>% 
  filter(year > plt_start_yr, 
         year <= plt_end_yr), 
       aes(x = year, y = scaled_yield)) + 
  facet_grid(reserve_frac ~ f_fmsy,
               labeller = labeller(f_fmsy = label_both,
                                   reserve_frac = label_both)) +
  geom_line() +
  geom_line(data = scaled_recs_plt %>% filter(lagged_year > plt_start_yr, 
         lagged_year <= plt_end_yr), aes(x = lagged_year, y = scaled_recruits), 
            colour = "blue") +
  geom_text(
      data = corr_vals,
      x = 1.01 * plt_start_yr ,
      y = 3,
      aes(label = paste0("Corr: ", round(corr, 3)))
  ) +
  ylab("Scaled yield (black) and lagged + scaled recruits (blue)") +
  xlab("Year") +
  theme_bw()

print(lagged_recs_yield_plt)

postscript(file = file.path(out_path, paste0(exp, "-lagged-recs-yield-ts-plt.eps")), 
           width = 8,
           heigh = 6)
print(lagged_recs_yield_plt)
invisible(dev.off())

png(
  file = file.path(out_path, paste0(exp, "-lagged-recs-yield-ts-plt.png")),
  width = 8,
  heigh = 6,
  units = "in", res = 600
)
print(lagged_recs_yield_plt)
invisible(dev.off())

```