### 2023-11-04 ms_figures_supp_figs.qmd
initial steps/prep results of simulations to create figs for manuscript.

---
title: "Recruits, abundance, and yield time series and contours"
author: "Botsford, Hastings, White, and Kilduff"
execute: 
  warning: false
format:
  html: 
    page-layout: full
    embed-resources: true
    anchor-sections: true
    fig-width: 7
    fig-height: 5
  pdf:
    fig-width: 7
    fig-height: 5
  docx:
    fig-width: 7
    fig-height: 5
editor: visual
params:
  experiment: "larval_pool_blue_rockfish_mei_135_75"
---

## Config and set up

```{r}
#| label: Good label name
#| echo: false
#| 
# c("larval_pool_blue_rockfish_mei_135_75",
# "larval_pool_blue_rockfish_white_135_75",
# "larval_pool_cabezon_mei_135_75",
# "larval_pool_cabezon_white_135_75",
# "larval_pool_china_rockfish_mei_135_75",
# "larval_pool_china_rockfish_white_135_75",
# "no_dispersal_blue_rockfish_mei_135_75",
# "no_dispersal_blue_rockfish_white_135_75",
# "no_dispersal_cabezon_mei_135_75",
# "no_dispersal_cabezon_white_135_75",
# "no_dispersal_china_rockfish_mei_135_75",
# "no_dispersal_china_rockfish_white_135_75")

main_figs_path <- file.path("output", "main_figs")
supp_figs_path <- file.path("output", "supp_figs")

if (!dir.exists(main_figs_path)) {
  dir.create(main_figs_path, recursive = TRUE) 
}

if (!dir.exists(supp_figs_path)) {
  dir.create(supp_figs_path, recursive = TRUE) 
}

con <- DBI::dbConnect(duckdb::duckdb(),
                      dbdir = "data/sim_out.duckdb",
                      read_only = FALSE)
```

```{r}
#| label: load-packages-data
#| echo: false

library(targets)
suppressMessages(library(tidyverse))
library(duckdb)
options(dplyr.summarise.inform = FALSE)
# Load relevant data from target objexts
tar_load(sim_nums)
tar_load(sim_len)
tar_load(species_flep_fished_ind)
tar_load(fishing_mortality_values)
tar_load(species_parms)
tar_load(sim_species_parms)
tar_load(sim_species_derived_vars)

# Load project functions
source("R/functions.R")

# constants
burn_in_parm <- 500
sim_len_parm <- 135

x_text_position_parm <- 550

waa_g <- sim_species_parms[[nice_species]][["biom_const"]] * sim_species_derived_vars[[nice_species]][["length_at_age"]] ^
                    sim_species_parms[[nice_species]][["biom_exp"]]

# use functions to create useful data structures
# FLEP to F reference tables by species
spec_fleps_2_fs <-
  get_fs_from_fleps(species = nice_species,
                    flep_inds = species_flep_fished_ind,
                    f_vals =  fishing_mortality_values)

```

## Main figures

### Figure 3

Contour plots showing the buffering effects of harvest rate and marine reserves on the abundance of a fished population, when there is no larval dispersal between the fished and reserve habitat patches.

Species: Blue rockfish

Dispersal: No Dispersal

Environmental noise: ENSO (MEI)

```{r}
#| label: figure_3

set.seed(11)
num_sims_2_avg <- 5
sims_2_use <- sort(sample(1:sim_nums, size = num_sims_2_avg, replace = FALSE))

table_name <- "no_dispersal_blue_rockfish_mei_135_75"

qry <- paste0("SELECT  sim_num, reserve_frac, flep_ratio, year, age, 
              sum(number) as abundance,
              sum(yield) as yield,
              FROM '", table_name, 
              "' WHERE sim_num IN (", toString(sims_2_use), ") AND year > 500 
              GROUP BY sim_num, reserve_frac, flep_ratio, year, age
              ORDER BY sim_num, reserve_frac, flep_ratio, year, age
              ")

out <- dbGetQuery(conn = con, qry)

out_2 <- out %>%
  group_by(sim_num, reserve_frac, flep_ratio) %>%
  summarize(abundance = mean(abundance, na.rm = TRUE),
            yield = mean(yield, na.rm=TRUE)) %>%
  ungroup()

ggplot(out_2, aes(x = as.numeric(reserve_frac), 
                y = as.numeric(flep_ratio),
                z = abundance)) +
  facet_grid(sim_num ~ .) +
  geom_contour_filled()

out_3 <- out_2 %>%
  group_by(reserve_frac, flep_ratio) %>%
  summarize(abundance = mean(abundance, na.rm = TRUE),
            yield = mean(yield, na.rm = TRUE)) %>%
  ungroup()

ggplot(out_3, aes(x = as.numeric(reserve_frac), 
                y = as.numeric(flep_ratio),
                z = abundance)) +
  geom_contour_filled()

# qry <- "SELECT reserve_frac, flep_ratio, sum(number) as total_nums, sum(yield) as total_yield from '", table_name, "'where year > 500 group by reserve_frac, flep_ratio order by reserve_frac asc, flep_ratio asc"

```

### Figure 4

Contour plots showing the buffering effects of harvest rate and marine reserves on the abundance of a fished population, when the fished and reserve patches share a larval pool.

Species: Blue rockfish

Dispersal: Larval pool

Environmental noise: ENSO (MEI)

```{r}
#| label: figure_4
```

### Figure 5

Representative timeseries of total population biomass under different combinations of fishing rate and marine reserve protection.

Species: Blue rockfish

Dispersal: Larval pool

Environmental noise: ENSO (MEI)

```{r}
#| label: figure_5

```

### Figure 6

Representative timeseries of fishery yield under different combinations of fishing rate and marine reserve protection.

Species: Blue rockfish

Dispersal: Larval pool

Environmental noise: ENSO (MEI)

```{r}
#| label: figure_6

```

### Figure 7

Representative timeseries of fishery yield (black curves) in the same harvest and marine reserve scenarios depicted in Fig. 5.

Species: Blue rockfish

Dispersal: Larval pool

Environmental noise: ENSO (MEI)

```{r}
#| label: figure_7

```